### 2016-11-16 Paper_v1.tex
Nov16

The ensemble of kinetic CFPS models captured the time evolution of CAT biosynthesis (Fig.~\ref{fig:CarbonBoth} - \ref{fig:Amino}).
The cell-free \textit{E.~coli} metabolic network was constructed by removing growth associated reactions from the iAF1260 reconstruction \cite{Feist:2007aa},
and by adding reactions describing chloramphenicol acetyltransferase (CAT) biosynthesis, a model protein for which we have a comprehensive training dataset \cite{2005_calhoun_BiotechnologyProgress}.
The CFPS model equations were formulated using the hybrid cell-free modeling framework of Wayman et al. \cite{pr3010138}.
An ensemble of model parameters (N $>$ 10,000) was estimated from measurements of glucose, CAT, organic acids (pyruvate, lactate, acetate, succinate, malate), energy species (A(x)P, G(x)P, C(x)P, U(x)P), and 18 of the 20 proteinogenic amino acids using a constrained Markov Chain Monte Carlo (MCMC) approach.
The MCMC algorithm minimized the error between the training data and model simulations starting from an initial parameter set assembled from literature and inspection.
Parameter sets were selected for the ensemble based upon their error, and the Pearson correlation coefficient between the candidate and existing sets the ensemble.
The parameter set with the lowest error value was defined as the best-fit set.
Central carbon metabolism (Fig.~\ref{fig:CarbonBoth}, top), energy species (Fig.~\ref{fig:Energy}), and amino acids (Fig.~\ref{fig:Amino}) were captured by the ensemble and the best-fit set.
The constrained MCMC approach estimated parameter sets with a median error greater than two-order of magnitude less than random parameter sets generated within
the same parameter bounds (Fig. \ref{fig:BoxPlot}); thus, we have confidence in the predictive capability of the estimated parameters.
Allosteric control was important to the dynamics of the organic acid intermediates and CAT biosynthesis (Fig.~\ref{fig:CarbonBoth}, bottom).
The acetate, lactate, pyruvate, succinate, malate and CAT trajectories were qualitatively different in the absence of allosteric control following glucose exhaustion.
In particular, the rate of CAT biosynthesis and lactate accumulation and subsequent consumption decreased following glucose exhaustion in the absence of control.

@@

To better understand which parameters, and parameter combinations influenced model performance we performed sensitivity analysis (Fig.~\ref{fig:Sensitivity}).
CAT production was most sensitive to the CAT synthesis reaction, oxidative phosphorylation activity, and alanine synthesis (Fig.~\ref{fig:Sensitivity}, top, section A).
The 16 next most important reactions to CAT production (section B) came from various pathways across the network: four each from glycolysis, the TCA cycle, and amino acid synthesis/degradation; two from the pentose phosphate pathway; and one each from the Entner-Doudoroff pathway and the energy species reactions.
The pairwise sensitivities (off-diagonal elements) were different from the corresponding first-order sensitivities (diagonal elements), and led to interesting outcomes.
For example, glutamine synthesis and arginine degradation were both among the most important reactions to CAT production (they rank 5th and 10th, respectively).
This was likely because they both affect the sensitive glutamine-glutamate balance; glutamine synthesis consumes glutamate, while arginine degradation produces it.
However, when both were perturbed, their combined effect on the model was low, as the respective contributions to consumption and production of glutamate cancelled.
The system state as a whole was most sensitive to glucose uptake via GTP and the forward reaction of lactate dehydrogenase (Fig.~\ref{fig:Sensitivity}, bottom, section F).
The 30 next most important reactions to the system state (section G) came from various pathways across the network: eight from amino acid synthesis/degradation; six from glycolysis; four from the TCA cycle; and two each from the pentose phosphate pathway, Entner-Doudoroff, energy/reducing pathways, and small molecule transport; one from oxidative phosphorylation; and one from pyrophosphatase consumption. The system state had more pairwise sensitivities that differ from the corresponding first-order sensitivities and stand out as significant.
For example, the first-order effect of alanine synthesis was large; it consumes both pyruvate and glutamate, two key species in the network.
In addition, a handful of alanine synthesis pairwise sensitivities were also large.
However, there were enough reactions that, when paired with alanine synthesis, had little effect on the model; malic enzyme is one of these, as it produces the pyruvate that alanine synthesis consumes. Thus, the total-order alanine synthesis sensitivity was low, placing it at the very bottom of section I.
Another interesting result was the intersection of sections F and G with section J.
The 53 reactions in section J were turned off in the best-fit set (rate constants were set to 0);
therefore, the perturbation of these reactions had no effect on the model.
Thus, all pairwise sensitivities with reactions in section J were pseudo first-order sensitivities for the other reactions.
Interestingly, many reactions in section F and several in section G showed their highest sensitivities when paired with the "non-effects" of section J.
Of these, three involved pyruvate, strengthening its role as a key metabolite; the others were glucose consumption via GTP/CTP-specific hexokinases,
fumarate reductase, and SO$_4$ utilization.
This suggested that these reactions' effects on the model were canceled out or lessened by most other reactions, but were of course not affected by the reactions in section J.
This was also likely the reason that reactions in section J rank above those in section K, despite having no effect themselves on the model.
Taken together, sensitivity analysis identified blocks of parameters that either individually, or in combination influenced model performance.
However, the sensitivity analysis did not establish what the maximum performance of the system was.
To answer that question we performed sequence-specific flux balance analysis of CAT production.

@@

\subsection*{Estimation of kinetic model parameters.}
We generated an ensemble of diverse parameter sets using a constrained Markov Chain Monte Carlo (MCMC) random walk strategy.
Starting from a single best fit parameter set estimated by inspection and literature,
we calculated the cost function, equal to the sum-squared-error between experimental data and model predictions:
\begin{equation}\label{eqn:cost-function}
    \texttt{cost}=\sum_{i=1}^{\mathcal{D}}\left[\frac{w_i}{\mathcal{Y}_{i}^2}\sum_{j=1}^{\mathcal{T}_i}\bigg(y_{ij}-x_{i}|_{t(j)}\bigg)^2 \right]
\end{equation}
where $\mathcal{D}$ denotes the number of datasets ($\mathcal{D}$~=~37),
$w_i$ denotes the weight of the $i^{th}$ dataset,
$\mathcal{T}_i$ denotes the number of timepoints in the $i^{th}$ dataset,
$t(j)$ denotes the $j^{th}$ timepoint, $y_{ij}$ denotes the measurement value of the $i^{th}$ dataset at the $j^{th}$ timepoint,
and $x_{i}|_{t(j)}$ denotes the simulated value of the metabolite corresponding to the $i^{th}$ dataset, interpolated to the $j^{th}$ timepoint.
Lastly, the cost calculation was scaled by the maximum experimental value in the $i^{th}$ dataset, $\mathcal{Y}_{i}=\max_{j}\left(y_{ij}\right)$.
We then perturbed each model parameter:
\begin{equation}\label{eqn:parameter-perturbation}
    k_i^{new}=k_i \cdot \exp(a\,r_i)\qquad{i=1,2,\hdots,\mathcal{P}}
\end{equation}
where $\mathcal{P}$ denotes the number of parameters ($\mathcal{P}$~=~815),
which includes 163 rate constants, 163 enzyme activity decay constants, 455 saturation constants, and 34 control parameters,
$k_i^{new}$ denotes the new value of the $i^{th}$ parameter,
$k_i$ denotes the current value of the $i^{th}$ parameter,
$a$ denotes a distribution variance, and $r_i$ denotes a random sample from the normal distribution.
For each newly generated parameter set, we re-solved the balance equations and calculated the cost function.
All sets with a lower cost than the previous set, and some with higher cost, were added to the ensemble.
After generating 12,437 sets, we selected 100 sets with minimal set to set correlation to avoid over-sampling any region of parameter space.
The original 12,437-set ensemble had a mean Pearson correlation coefficient of 0.94 between pairs of sets,
while the 100-set ensemble had a mean Pearson correlation coefficient of 0.83 between pairs of sets.

@@

\subsection*{Sequence-specific FBA and calculation of CAT yield}
The yield on CAT production was calculated for each case as a ratio of carbon produced as CAT to carbon consumed as reactants (glucose and amino acids):
\begin{equation}\label{eqn:yield-definition}
	Yield=\frac{\Delta m_{CAT}\;C_{CAT}}{\sum_{i=1}^{\mathcal{R}}\max(\Delta m_{i},0)\;C_{m_i}}
\end{equation}
where $\Delta m_{CAT}$ denotes the amount of CAT produced, $C_{CAT}$ denotes carbon number of CAT, $\mathcal{R}$ denotes the number of reactants, $\Delta m_{i}$ denotes the amount of the $i^{th}$ reactant consumed, never allowed to be negative, and $C_{m_i}$ denotes the carbon number of the $i^{th}$ reactant.
Because no data was available for arginine or glutamate, these reactants were left out of the experimental yield calculation.
Yield of the best-fit parameter set and the experimental data were calculated by setting $\Delta CAT$ equal to the final minus the initial CAT concentration and setting $\Delta m_{i}$ equal to the initial minus the final reactant concentration.
Theoretical maximum CAT carbon yields for three cases discussed below were calculated using flux balance anaylsis (FBA) with a sequence-specific description of CAT synthesis, where $\Delta m_{i}$ denotes the flux of the $i^{th}$ species.
This sequence-specific FBA \citep{Allen:2003aa} problem was formulated as:
\begin{equation}\nonumber
 \begin{multlined}
	\qquad \qquad \qquad \max_{\boldsymbol{w}}{} \! \left( w_{obj} = \mathbf{\boldsymbol{\theta}}^T \boldsymbol{w} \right) \\
	\mathrm{Subject \; to:}
	 \; \; \mathbf{S}\mathbf{w}=\mathbf{0} \\
\alpha_i \leq w_i \leq \beta_i  \qquad i=1,2,\hdots,\mathcal{R}
 \end{multlined}
\end{equation}
where $\mathbf{S}$ denotes the stoichiometric matrix, $\mathbf{w}$ denotes the unknown flux vector, $\boldsymbol{\theta}$ denotes the objective selection vector
and $\alpha_i$ and $\beta_i$ denote the lower and upper bounds on flux $w_{i}$, respectively.
The objective $w_{obj}$ was to maximize the specific rate of CAT formation.
The specific glucose uptake rate was constrained to allow a maximum flux of 40 mM/hr according to experimental data; the specific amino acid uptake rates were bound to allow a maximum flux of 30 mM/hr, but did not reach this maximum flux.
The transcription and translation template reactions come from sequence-specific analysis \cite{Allen:2003aa}, and include transcription initiation, transcription, mRNA degradation, translation initiation, translation, and tRNA charging.
The flux balance analysis problem was solved using the GNU Linear Programming Kit (v4.52) \cite{GLPK}.
The solution flux vector was used to calculate the carbon yield of CAT for the three FBA cases.
Glucose, oxygen, and amino acids were modeled as being imported into the system, while CAT synthesis and metabolite byproduct formation were modeled as export from the system.
The rest of the network followed a pseudo steady-state assumption where metabolites were not allowed to accumulate; thus, the network could be solved by linear programming instead of solving differential equations.

@@

An ensemble of 100 sets of flux distributions was calculated for three different cases: unconstrained, constrained by transcription and translation (TX/TL) rates, and  constrained by TX/TL rates and experimental data.
For the unconstrained case, all rates were left unbounded, except the specific glucose uptake rate.
An ensemble of flux distributions was then calculated by randomly sampling the maximum specific glucose uptake rate from within a range of 30 to 40 mM/hr, determined from experimental data.
For the case constrained by TX/TL rates, an ensemble was generated by randomly sampling RNAP polymerase levels, ribosome levels, and elongation rates in a physiological range determined from literature.
RNA polymerase levels were sampled between 60 and 80 nM, ribosome levels between 7 and 16 \textmu M, the RNA polymerase elongation rate between 20 and 30 nt/sec, and the ribosome elongation rate between 1.5 and 3 aa/sec \cite{2005_underwood_biotech, Garamella:2016aa}.
For the case constrained by TX/TL rates and experimental data, the lower and upper bounds on the fluxes for the data-informed metabolites were sampled within the range given by the experimental noise.
This included the data for glucose, organic acids, energy species, and amino acids; CAT was not constrained by experimental data, but by the TX/TL rates as stated above.

>>>>

The ensemble of kinetic CFPS models captured the time evolution of CAT biosynthesis (Fig.~\ref{fig:CarbonBoth} - \ref{fig:Amino}).
The cell-free \textit{E.~coli} metabolic network was constructed by removing growth associated reactions from the iAF1260 reconstruction \cite{Feist:2007aa},
and by adding reactions describing chloramphenicol acetyltransferase (CAT) biosynthesis, a model protein for which we have a comprehensive training dataset \cite{2005_calhoun_BiotechnologyProgress}.
The CFPS model equations were formulated using the hybrid cell-free modeling framework of Wayman et al. \cite{pr3010138}.
An ensemble of model parameters (N $>$ 10,000) was estimated from measurements of glucose, CAT, organic acids (pyruvate, lactate, acetate, succinate, malate), energy species (A(x)P, G(x)P, C(x)P, U(x)P), and 18 of the 20 proteinogenic amino acids using a constrained Markov Chain Monte Carlo (MCMC) approach.
The MCMC algorithm minimized the error between the training data and model simulations starting from an initial parameter set assembled from literature and inspection.
Parameter sets were selected for the ensemble based upon their error, and the Pearson correlation coefficient between the candidate and existing sets in the ensemble.
The parameter set with the lowest error value was defined as the best-fit set.
Central carbon metabolism (Fig.~\ref{fig:CarbonBoth}, top), energy species (Fig.~\ref{fig:Energy}), and amino acids (Fig.~\ref{fig:Amino}) were captured by the ensemble and the best-fit set.
The constrained MCMC approach estimated parameter sets with a median error greater than two-order of magnitude less than random parameter sets generated within
the same parameter bounds (Fig. \ref{fig:BoxPlot}); thus, we have confidence in the predictive capability of the estimated parameters.
Allosteric control was important to the dynamics of the organic acid intermediates and CAT biosynthesis (Fig.~\ref{fig:CarbonBoth}, bottom).
The acetate, lactate, pyruvate, succinate, malate and CAT trajectories were qualitatively different in the absence of allosteric control following glucose exhaustion.
In particular, the rate of CAT biosynthesis and lactate accumulation and subsequent consumption decreased following glucose exhaustion in the absence of control.

@@

To better understand which parameters and parameter combinations influenced model performance we performed sensitivity analysis (Fig.~\ref{fig:Sensitivity}).
CAT production was most sensitive to the CAT synthesis reaction, oxidative phosphorylation activity, and alanine synthesis (Fig.~\ref{fig:Sensitivity}, top, section A).
The 16 next most important reactions to CAT production (section B) came from various pathways across the network: four each from glycolysis, the TCA cycle, and amino acid synthesis/degradation; two from the pentose phosphate pathway; and one each from the Entner-Doudoroff pathway and the energy species reactions.
The pairwise sensitivities (off-diagonal elements) were different from the corresponding first-order sensitivities (diagonal elements), and led to interesting outcomes.
For example, glutamine synthesis and arginine degradation were both among the most important reactions to CAT production (they rank 5th and 10th, respectively).
This was likely because they both affect the sensitive glutamine-glutamate balance; glutamine synthesis consumes glutamate, while arginine degradation produces it.
However, when both were perturbed, their combined effect on the model was low, as the respective contributions to consumption and production of glutamate cancelled.
The system state as a whole was most sensitive to glucose uptake via GTP and the forward reaction of lactate dehydrogenase (Fig.~\ref{fig:Sensitivity}, bottom, section F).
The 30 next most important reactions to the system state (section G) came from various pathways across the network: eight from amino acid synthesis/degradation; six from glycolysis; four from the TCA cycle; and two each from the pentose phosphate pathway, Entner-Doudoroff, energy/reducing pathways, and small molecule transport; one from oxidative phosphorylation; and one from pyrophosphatase consumption. The system state had more pairwise sensitivities that differ from the corresponding first-order sensitivities and stand out as significant.
For example, the first-order effect of alanine synthesis was large; it consumes both pyruvate and glutamate, two key species in the network.
In addition, a handful of alanine synthesis pairwise sensitivities were also large.
However, there were enough reactions that, when paired with alanine synthesis, had little effect on the model; malic enzyme is one of these, as it produces the pyruvate that alanine synthesis consumes. Thus, the total-order alanine synthesis sensitivity was low, placing it at the very bottom of section I.
Another interesting result was the intersection of sections F and G with section J.
The 53 reactions in section J were turned off in the best-fit set (rate constants were set to 0);
therefore, the perturbation of these reactions had no effect on the model.
Thus, all pairwise sensitivities with reactions in section J were pseudo first-order sensitivities for the other reactions.
Interestingly, many reactions in section F and several in section G showed their highest sensitivities when paired with the "non-effects" of section J.
Of these, three involved pyruvate, strengthening its role as a key metabolite; the others were glucose consumption via GTP/CTP-specific hexokinases,
fumarate reductase, and SO$_4$ utilization.
This suggested that these reactions' effects on the model were canceled out or lessened by most other reactions, but were of course not affected by the reactions in section J.
This was also likely the reason that reactions in section J rank above those in section K, despite having no effect themselves on the model.
Taken together, sensitivity analysis identified blocks of parameters that either individually, or in combination influenced model performance.
However, the sensitivity analysis did not establish what the maximum performance of the system was.
To answer that question we performed sequence-specific flux balance analysis of CAT production.

@@

\subsection*{Estimation of kinetic model parameters.}
We generated an ensemble of diverse parameter sets using a constrained Markov Chain Monte Carlo (MCMC) random walk strategy.
Starting from a single best fit parameter set estimated by inspection and literature,
we calculated the cost function, equal to the sum-squared-error between experimental data and model predictions:
\begin{equation}\label{eqn:cost-function}
    \texttt{cost}=\sum_{i=1}^{\mathcal{D}}\left[\frac{w_i}{\mathcal{Y}_{i}^2}\sum_{j=1}^{\mathcal{T}_i}\bigg(y_{ij}-x_{i}|_{t(j)}\bigg)^2 \right]
\end{equation}
where $\mathcal{D}$ denotes the number of datasets ($\mathcal{D}$~=~37),
$w_i$ denotes the weight of the $i^{th}$ dataset,
$\mathcal{T}_i$ denotes the number of timepoints in the $i^{th}$ dataset,
$t(j)$ denotes the $j^{th}$ timepoint, $y_{ij}$ denotes the measurement value of the $i^{th}$ dataset at the $j^{th}$ timepoint,
and $x_{i}|_{t(j)}$ denotes the simulated value of the metabolite corresponding to the $i^{th}$ dataset, interpolated to the $j^{th}$ timepoint.
Lastly, the cost calculation was scaled by the maximum experimental value in the $i^{th}$ dataset, $\mathcal{Y}_{i}=\max_{j}\left(y_{ij}\right)$.
We then perturbed each model parameter:
\begin{equation}\label{eqn:parameter-perturbation}
    k_i^{new}=k_i \cdot \exp(a\,r_i)\qquad{i=1,2,\hdots,\mathcal{P}}
\end{equation}
where $\mathcal{P}$ denotes the number of parameters ($\mathcal{P}$~=~815), which includes 163 rate constants, 163 enzyme activity decay constants, 455 saturation constants, and 34 control parameters, $k_i^{new}$ denotes the new value of the $i^{th}$ parameter, $k_i$ denotes the current value of the $i^{th}$ parameter, $a$ denotes a distribution variance, and $r_i$ denotes a random sample from the normal distribution.
Parameters were kept between an upper and lower bound that varied by parameter type:
\begin{equation}\label{eqn:parameter-bounds}
    k_i^{new,bounded}=\min\left(\max\left(k_i^{new},l_i\right),u_i\right)
\end{equation}
where $k_i^{new,bounded}$ denotes the bounded parameter value, $l_i$ denotes the lower bound for that parameter type, and $u_i$ denotes the upper bound for that parameter type.
Rate constants were bounded between 0 and 500000 mM/h [REFERENCE NEEDED]; this resulted in a mean of 324.6 mM/h and a standard deviation of 4311.2 mM/h for rate constants, giving a coefficient of variation (CV) of 13.3.
Enzyme degradation rates were bounded between 0 and 1 h$^{-1}$ [REFERENCE NEEDED]; $\mu$ = 0.19 h$^{-1}$, $\sigma$ = 0.32 h$^{-1}$, CV = 1.64.
Saturation constants were bounded between 0.001 and 10 mM [REFERENCE NEEDED]; $\mu$ = 0.76 mM, $\sigma$ = 1.71 mM, CV = 2.25.
Control parameters (gains and orders) were left unbounded.
Control gains: $\mu$ = 0.51, $\sigma$ = 1.21, CV = 2.35.
Control orders: $\mu$ = 14.6, $\sigma$ = 66.0, CV = 4.53.
For each newly generated parameter set, we re-solved the balance equations and calculated the cost function.
All sets with a lower cost than the previous set, and some with higher cost, were added to the ensemble.
After generating 12,437 sets, we selected 100 sets with minimal set to set correlation to avoid over-sampling any region of parameter space.
The original 12,437-set ensemble had a mean Pearson correlation coefficient of 0.94 between pairs of sets, while the 100-set ensemble had a mean Pearson correlation coefficient of 0.83 between pairs of sets.

@@

\subsection*{Sequence-specific FBA and calculation of CAT yield}
The yield on CAT production was calculated for each case as a ratio of carbon produced as CAT to carbon consumed as reactants (glucose and amino acids):
\begin{equation}\label{eqn:yield-definition}
	Yield=\frac{\Delta m_{CAT}\;C_{CAT}}{\sum_{i=1}^{\mathcal{R}}\max(\Delta m_{i},0)\;C_{m_i}}
\end{equation}
where $\Delta m_{CAT}$ denotes the amount of CAT produced, $C_{CAT}$ denotes carbon number of CAT, $\mathcal{R}$ denotes the number of reactants, $\Delta m_{i}$ denotes the amount of the $i^{th}$ reactant consumed, never allowed to be negative, and $C_{m_i}$ denotes the carbon number of the $i^{th}$ reactant.
Because no data was available for arginine or glutamate, these reactants were left out of the experimental yield calculation.
Yield of the best-fit parameter set and the experimental data were calculated by setting $\Delta CAT$ equal to the final minus the initial CAT concentration and setting $\Delta m_{i}$ equal to the initial minus the final reactant concentration.
Theoretical maximum CAT carbon yields for three cases discussed below were calculated using flux balance anaylsis (FBA) with a sequence-specific description of CAT synthesis, where $\Delta m_{i}$ denotes the flux of the $i^{th}$ species.
This sequence-specific FBA \citep{Allen:2003aa} problem was formulated as:
\begin{equation}\nonumber
 \begin{multlined}
	\qquad \qquad \qquad \max_{\boldsymbol{w}}{} \! \left( w_{obj} = \mathbf{\boldsymbol{\theta}}^T \boldsymbol{w} \right) \\
	\mathrm{Subject \; to:}
	 \; \; \mathbf{S}\mathbf{w}=\mathbf{0} \\
\alpha_i \leq w_i \leq \beta_i  \qquad i=1,2,\hdots,\mathcal{R}
 \end{multlined}
\end{equation}
where $\mathbf{S}$ denotes the stoichiometric matrix, $\mathbf{w}$ denotes the unknown flux vector, $\boldsymbol{\theta}$ denotes the objective selection vector
and $\alpha_i$ and $\beta_i$ denote the lower and upper bounds on flux $w_{i}$, respectively.
The objective $w_{obj}$ was to maximize the specific rate of CAT formation.
The specific glucose uptake rate was constrained to allow a maximum flux of 40 mM/h according to experimental data; the specific amino acid uptake rates were bound to allow a maximum flux of 30 mM/h, but did not reach this maximum flux.
The transcription and translation template reactions come from sequence-specific analysis \cite{Allen:2003aa}, and include transcription initiation, transcription, mRNA degradation, translation initiation, translation, and tRNA charging.
The flux balance analysis problem was solved using the GNU Linear Programming Kit (v4.52) \cite{GLPK}.
The solution flux vector was used to calculate the carbon yield of CAT for the three FBA cases.
Glucose, oxygen, and amino acids were modeled as being imported into the system, while CAT synthesis and metabolite byproduct formation were modeled as export from the system.
The rest of the network followed a pseudo steady-state assumption where metabolites were not allowed to accumulate; thus, the network could be solved by linear programming instead of solving differential equations.

@@

An ensemble of 100 sets of flux distributions was calculated for three different cases: unconstrained, constrained by transcription and translation (TX/TL) rates, and  constrained by TX/TL rates and experimental data.
For the unconstrained case, all rates were left unbounded, except the specific glucose uptake rate.
An ensemble of flux distributions was then calculated by randomly sampling the maximum specific glucose uptake rate from within a range of 30 to 40 mM/h, determined from experimental data.
For the case constrained by TX/TL rates, an ensemble was generated by randomly sampling RNAP polymerase levels, ribosome levels, and elongation rates in a physiological range determined from literature.
RNA polymerase levels were sampled between 60 and 80 nM, ribosome levels between 7 and 16 \textmu M, the RNA polymerase elongation rate between 20 and 30 nt/sec, and the ribosome elongation rate between 1.5 and 3 aa/sec \cite{2005_underwood_biotech, Garamella:2016aa}.
For the case constrained by TX/TL rates and experimental data, the lower and upper bounds on the fluxes for the data-informed metabolites were sampled within the range given by the experimental noise.
This included the data for glucose, organic acids, energy species, and amino acids; CAT was not constrained by experimental data, but by the TX/TL rates as stated above.
